---
title: "Testing "
output: html_notebook
---

# Experimental Setup
In order to test the different implementations of focal window filtering using `raster` or `terra`, a sample file is processed. It may be important to note though, here only the actual packages are compared. With respect to `geoprocessors`, it would be useful to test the eventual implementation been called from a comandline as well as the GeoTIFFs actual file properties.
```{r}
opt <- list(
  input = system.file("ex/logo.tif", package="terra"),
  output = NULL,  # Not writing to file in this test
  radius = 10,
  fun = "mean",
  overwrite = FALSE
)
```


# RASTER Version
First, the initial implementation by `mwip/geoprocessors` is run (commit #3303fd2).
```{r}
library(raster)

raster_implementation <- function(opt) {
  # load input image
  r <- raster(opt$input)
  
  # create circular focal weight matrix
  w <- focalWeight(r, d = opt$radius, type = 'circle')
  w <- ifelse(w > 0, 1, NA)
  
  the_fun <- get(opt$fun)
  
  # apply focal matrix ver image
  focal(r, w,
        fun = function(x, na.rm, ... = the_fun) {
          if (na.rm) {
            x <- na.omit(x)
          }
          ...(x)
        },
        pad = TRUE, pad.value = NA, na.rm = TRUE)
}

# Visualize Result
res_raster <- raster_implementation(opt)
plot(res_raster)

```

# TERRA Version
Analog, the sample file is processed with the `terra`-package next. Be aware though, it's syntax is slightly different. That said parameters such as `pad = TRUE` and `pad.value = NA` both are absent.
```{r}
library(terra)

terra_implementation <- function(opt) {
  # load input image
  r <- rast(opt$input)
  
  # create circular focal weight matrix
  w <- focalMat(r, d = opt$radius, type = 'circle')
  w <- ifelse(w > 0, 1, NA)
  
  the_fun <- get(opt$fun)
  
  # apply focal matrix ver image
  focal(r, w,
        fun = function(x, na.rm, ... = the_fun) {
          if (na.rm) {
            x <- na.omit(x)
          }
          ...(x)
        },
        na.rm = TRUE)
}
res_terra <- terra_implementation(opt)

# Visualize Result
plot(res_terra)
```

# Compare Results
When comparing the results, it can be seen that both packages come with their individual structures. The digital numbers though are identical.
```{r}
str(res_raster)
str(res_terra)

values_raster <- as.numeric(res_raster[])
values_terra <- as.numeric(res_terra$red[])
identical(values_raster, values_terra)
```

# Benchmark
Last but not least, a benchmark is run in order to see computational benefits of using `terra`. Note, it would be interesting though to expand this benchmark with growing window sizes...
```{r}
library(microbenchmark)
microbenchmark(
  raster = raster_implementation(opt),
  terra  = terra_implementation(opt)
)
boxplot(.Last.value)
```

